---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# stdcab

<!-- badges: start -->
[![R-CMD-check](https://github.com/suvedimukti/stdcab/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/suvedimukti/stdcab/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of the **stdcab** package is to apply spatial thinning on multi-class spatial point data, spatial cluster analysis for random and repeated cross-validations which partitions data into training and testing set compatible to tidy-modeling resamples. The package allows to run semivariance analysis, plotting and spatial blocking of observations.

# Installation

You can install the released version of stdcab from [CRAN](https://cran.r-project.org) with:

``` {r 
install.packages("stdcab")
```

To install the latest stdcab version from [GitHub](https://github.com/) use:

```{r gitinstall}
remotes::install_github('suvedimukti/stdcab')
```
 Repeated Cluster CV: `sample` compatible

 This function is extended based on `spatial_cluster_sample`, which is a wrapper around `spatial_clustering_cv` function of **sptialsample** package.
 
##  Load dataset

```{r example}
# load dataset
library(stdcab)
data(landcover)

# create another dataset based on landcover data that comes with the package
#
dspatial <- landcover

# examine data
head(dspatial)

# Class_name is the dependent (response ) data with seven classes (1 through  7)
```

## Visualize the data 

```{r datavis}
# load ggplot 2 for visualization
library(ggplot2)
ggplot(data = dspatial)+
  geom_sf(aes(colour = factor(Class_name)), size = 1.5, alpha = 0.8)+
  scale_colour_viridis_d()+
  labs(color = "Land Cover Class")+
  theme_bw(12)
```

## Apply repeated cluster sampling on sf data

To make visualization easy lets make five folds and five repeats resulting 25
splits of data based on **kmeans** clustering. 

```{r repsamp}
# setting seeds
set.seed(1318)

spc_rcv <- repeated_spatial_cluster_sample(data = dspatial, v = 5, repeats = 5,
                            coords = NULL, spatial = TRUE, clust_method = "kmeans", 
                            dist_clust = NULL)

spc_rcv


```

## Visualize clusters 

Following chunk of code is a function to run each split at a time to visualize 
`Analysis\Training` and `Assessment\Testing` set in each fold and repeats.

```{r clusvis1}
library(magrittr) # 
fplot_splits <- function(split) {
  gp <- analysis(split) %>%
   dplyr::mutate(analysis = "Training") %>%
   dplyr::bind_rows(assessment(split) %>%
   dplyr::mutate(analysis = "Testing")) %>%
    ggplot(aes(X, Y, color = analysis)) +
    geom_point(alpha = 0.7, size = 2) +
    coord_fixed() +
    labs(color = "Train/Test") +
    scale_color_viridis_d(direction = -1) +
    xlab("Longitude (m)") +
    ylab("Latitidue (m)")
  print(gp)
}
```

## Plotting 
```{r hclust_plot1, echo = TRUE,  animation.hook='gifski', eval = requireNamespace('gifski', quietly = TRUE)}
# plot using walk function from purrr package
# this requires gifski package
purrr::walk(spc_rcv$splits, fplot_splits)
```
