---
title: "Spatial Clustering of Point Data For Machine Learning in Tidy-modeling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial_Clustering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stdcab)
```

# Introduction

Ecological data usually have inherent spatial structures; nearer things are 
more similar than distant things (Toblers, 1970). This relationship can exist 
at any spatial scale from local to global with a positive spatial autocorrelation
as a nonzero covariance between spatially proximal observations or proximal values.  

This definition imposes the need for a means to measure, 
based on a sample of data values, the covariance between nearby points and to 
decide whether or not this covariance is consistent with a random spatial 
arrangement of values. Such non-randomness in the data are not problematic. Correlated data allows
to uncover patterns from the process. However, in parametric statistical analysis of 
data problems can arise 1) non-randomness of error; 2) pseudo-replication-if too many observations are
made within the distance at which observations are spatially structured, and 3) As the predictor variables
(Independent or explanatory variables) are often correlated with dependence structures, resulting model 
can over-inflate the model's accuracy by wholly or partially removing residual structures.

Parametric models can address dependence structures in the data (e.g., autoregressive or mixed model), at least in theory. 
However, in practice, model specification bias together with structural over-fitting, can seriously impair the 
diagnostic evaluations of the model. Further, well-known popular machine learning models (e.g., random forest)
don't allow accounting for such spatial dependence structures. 

## What happens with non-parametric models?

In an ideal model building ( or machine learning), evaluation and validation should be 
performed using independent data. For example, model evaluation data should not come from 
within the same geographic extent from which the model was built (spatially distinct). 
Two scenarios may arise in such cases: 1) either no such independent data exist; usually, 
we tend to focus on collecting data within the spatial extent of our study area. 
In classification problems of remote sensing data, we may collect sample data from on-screen (heads-up) digitization. 
We can't guarantee that the relationship in the data is still the same. 
2) Data set on hand may not meet the assumption of independence.

The current paradigm of evaluation predictive error of machine learning model(s) is 
through cross-validation, usually with the use of hold-out(saving some percent of 
total sample data). The key idea is that the training data is independent of hold-out 
(validation) data. However, training points tend to cluster due to resources or 
time limitations. While it is possible to create independent data apriori, this 
is not usually the case; because we can't be certain independence of data as 
spatial dependence can happen at any scale.

Hold-out data, however, does not necessarily get rid of the problem at hand, 
as testing/validation data can be drawn from a nearby location with a dependence 
structure. Therefore, the resulting model would favor the complex, and 
accuracy metrics would be too optimistic.


# How to deal with it





