---
title: "Spatial Grid/Block Sampling For TidyModelinng"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial Grid/Block Sampling For TidyModelinng}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stdcab)
```

# Introduction

In spatial analysis, creation of regular polygons within the extent of certain 
geographic data is common task. This could be done to create regular equal size 
grids. This could be useful in uniform sampling or retroactively random selection.

Spatial tessellation or grids splits the extent into series of sub-extent, which is then used to summarize the 
information at hand by that extent. For classification, to reduce the spatial structures of the data
the extent of the training data or study area can be divided into sub-extent/sub-units. The information that falls within 
the sub-extent then can be summarize to serve the purpose. For example, the frequency of criminal activities across county, number of trees in each 10 km grids etc.

How large the sub-units/sub-extent should be can be calculated using semi-variogram analysis. The goal here is to maximize the 
heterogeneity between blocks. This spatial grids is loosely similar to randomized control design in lot of aspects. 

The function `spatial_grid_sample` is similar to generate tessellations or create Fishnet tool in ArcGIS Pro or 
ArcGIS desktop software from Environmental System Research Institute (ESRI). This function only supports rectangles or squares.
The width and height information can be obtained from running `fit_variogram` or `multiple_variogram` functions. 


# Create spatial grid of 10km by 10km

To create the spatial grids the unit should be in the projection system, preferably in Universal 
Transverse Mercator or State Plane System. The function also allows to specify if groups of sub-grids needs to be selected as in the case of k-fold cross validation. For selection of sub-grids.

```{r gengrids, fig.height=8, fig.width=8}

data(landcover)

sp_grid<- spatial_grid_sample(data = landcover,cellsize = c(10000, 10000),
                              show_grid = TRUE,fold_selection = "default",
                              k = 1)

```

The `default` option ignores the value of `k` while splitting in the data, 
however expect the value. Other options are `random` selections of 
sub-grids into `k` groups, and systematic selection. 

If `show_grid` option is NULL, the the output will not have a grid-map. 
However, saves created grids and observations within each grid. 


```{r plotting_es, fig.height=8, fig.width=8}
#-----------------------#
sblock<- sp_grid$blocks

gp2 <- ggplot2::ggplot() +
      ggplot2::geom_sf(
        data = sblock, color = "purple",
        fill = "orangered4",
        alpha = 0.04,
        size = 0.7
      ) 
gp2
```


# Spatial grid resamples for Tidy-modeling

The final output from the data is the `splits` tibble. With a default methods, resultant grids
can be used as a leave-one-out cross validation. 

```{r rsplit, paged.print=TRUE}
# Create spatial grid of 10km by 10km
# make manual_rset object
msplits<- rsample::manual_rset(splits = sp_grid$splits$splits, ids = sp_grid$splits$id)

# convert to caert compatible format
caret_train_test<- rsample::rsample2caret(msplits,c("analysis","assessment"))

caret_train_test[1]$index$Fold01

```

In the above example, total 35 grids of 10km width are returned. Which are `rsample` compatible. 
In case these information needs to be transferred to `caret` compatible format, this can be easily accomplished
by using


# References @

Legendre, P., 1993. Spatial autocorrelation: problem or new paradigm? Ecology 74, 1659–1673.
Legendre, P., Dale, M.R.T., Fortin, M.-J., Gurevitch, J., Hohn, M., Myers, D., 2002. The consequences of spatial structure for the design and analysis of ecological field surveys. Ecography 25, 601–615.

Legendre, P., Fortin, M.J., 1989. Spatial pattern and ecological analysis.
Miller, J., Franklin, J., Aspinall, R., 2007. Incorporating spatial dependence in predictive vegetation models. Ecol. Modell. 202, 225–242. https://doi.org/10.1016/j.ecolmodel.2006.12.012

Miller, J.R., Turner, M.G., Smithwick, E.A.H., Dent, C.L., Stanley, E.H., 2004. Spatial extrapolation: the science of predicting ecological patterns and processes. BioScience 54, 310–320.

@ not an exhaustive list of references
